Reactive SINDy: Discovering governing reactions from concentration data

### 2018-12-04 docs/manuscript_jcp.tex
run mapk cv again with more points, plot results for mapk and predator-prey, wip manuscript

The goal is to find coefficients $\Xi=\begin{pmatrix}\xi_{1} & \xi_{2} & \cdots & \xi_{R}\end{pmatrix}^{\top}$,
so that 
\begin{align}
\dot{\textbf{X}}=\Theta(\textbf{X})\Xi=\sum_{r=1}^{R}\theta_{r}(\textbf{X})\xi_{r}.
\end{align}
In particular, the system is linear in the coefficients $\Xi$, which
makes regression tools such as elastic net regularization~\cite{Zou2005}
applicable. To this end, one can consider the regularized minimization
problem (reactive SINDy): 
\begin{align}
\hat{\Xi}&=\underset{\Xi}{\arg\min}\Bigg(\frac{1}{2T}\left\Vert \dot{\textbf{X}}-\Theta(\textbf{X})\Xi\right\Vert _{F}^{2} \label{method:minimizationproblem} \\
&\quad+\alpha\lambda\|\Xi\|_{1}+\alpha(1-\lambda)\|\Xi\|_{2}^{2}\Bigg)\quad\text{subject to }\Xi\geq0.\nonumber
\end{align}
Here, $\|\cdot\|_{F}$ denotes the Frobenius norm, $\lambda\in[0,1]$
is a hyperparameter that interpolates linearly between LASSO~\cite{Tibshirani1996,Hastie2009}
and Ridge~\cite{Hoerl1} methods, and $\alpha\geq0$ is a hyperparameter
that, depending on $\lambda$, can induce sparsity and give preference
to smaller solutions in the $L_{1}$ or $L_{2}$ sense. For $\alpha=0$
the minimization problem reduces to standard least-squares (LSQ) with
the constraint $\Xi\geq0$. Reactive SINDy (\ref{method:minimizationproblem})
is therefore a generalization of the SINDy method to to the vector-valued
ansatz functions.

@@

@@

\end{tabular}\caption{\label{tab:reaction-library}Full set of ansatz reactions $\Theta$
used in Sec.~\ref{sec:results}. The given rate constants define
the ground truth reaction model.}

@@

>>>>

The goal is to find coefficients $\Xi=\begin{pmatrix}\xi_{1} & \xi_{2} & \cdots & \xi_{R}\end{pmatrix}^{\top}$,
so that 
\begin{align}
\dot{\textbf{X}}=\Theta(\textbf{X})\Xi=\sum_{r=1}^{R}\theta_{r}(\textbf{X})\xi_{r}.
\end{align}
In particular, the system is linear in the coefficients $\Xi$, which
makes regression tools such as elastic net regularization~\cite{Zou2005}
applicable. To this end, one can consider the regularized minimization
problem (reactive SINDy): 
\begin{align}
\hat{\Xi}&=\underset{\Xi}{\arg\min}\Bigg(\frac{1}{2T}\left\Vert \dot{\textbf{X}}-\Theta(\textbf{X})\Xi\right\Vert _{F}^{2} \label{method:minimizationproblem} \\
&\quad+\alpha\lambda\|\Xi\|_{1}+\alpha(1-\lambda)\|\Xi\|_{2}^{2}\Bigg)\quad\text{subject to }\Xi\geq0.\nonumber
\end{align}
Here, $\|\cdot\|_{F}$ denotes the Frobenius norm, $\lambda\in[0,1]$
is a hyperparameter that interpolates linearly between LASSO~\cite{Tibshirani1996,Hastie2009}
and Ridge~\cite{Hoerl1} methods, and $\alpha\geq0$ is a hyperparameter
that, depending on $\lambda$, can induce sparsity and give preference
to smaller solutions in the $L_{1}$ or $L_{2}$ sense. For $\alpha=0$
the minimization problem reduces to standard least-squares (LSQ) with
the constraint $\Xi\geq0$. Reactive SINDy (\ref{method:minimizationproblem})
is therefore a generalization of the SINDy method to vector-valued
ansatz functions.

@@

\subsection{Application to MAPK cascade\label{sec:mapk}}

\subsection{Application to Lotka-Volterra system\label{sec:lv}}

@@

\end{tabular}\caption{\label{tab:reaction-library}Full set of ansatz reactions $\Theta$ used in Sec.~\ref{sec:results} for the gene-regulatory network. The given rate constants define the ground truth reaction model.}

@@

\begin{table*}[h!]
    \centering %
    %\resizebox{.9\textwidth}{!}{%
    \begin{tabular}{rclcc}
        \toprule 
        \multicolumn{3}{c}{Reaction} & rate  & description \tabularnewline
        \midrule 
        $\mathrm{DNA}_{\mathrm{A}}$  & $\rightharpoonup$  & $\mathrm{DNA}_{\mathrm{A}}+\mathrm{mRNA}_{\mathrm{A}}$  & $k_{1}=1.8$  & transcription of $\mathrm{mRNA}_{\mathrm{A}}$\tabularnewline
        $\mathrm{mRNA}_{\mathrm{A}}+\mathrm{A}$  & $\rightharpoonup$  & $\mathrm{mRNA}_{\mathrm{A}}$  & $k_{19}=0$  & artificial fusion\tabularnewline
        $\mathrm{mRNA}_{\mathrm{B}}+\mathrm{B}$  & $\rightharpoonup$  & $\mathrm{mRNA}_{\mathrm{B}}$  & $k_{20}=0$  & artificial fusion\tabularnewline
        \bottomrule
    \end{tabular}\caption{\label{tab:reaction-library-mapk}Full set of ansatz reactions $\Theta$ used in Sec.~\ref{sec:mapk} for the MAPK system. The given rate constants define the ground truth reaction model.}
\end{table*}

\begin{table*}[h!]
    \centering %
    %\resizebox{.9\textwidth}{!}{%
    \begin{tabular}{rclcc}
        \toprule 
        \multicolumn{3}{c}{Reaction} & rate  & description \tabularnewline
        \midrule 
        $\mathrm{DNA}_{\mathrm{A}}$  & $\rightharpoonup$  & $\mathrm{DNA}_{\mathrm{A}}+\mathrm{mRNA}_{\mathrm{A}}$  & $k_{1}=1.8$  & transcription of $\mathrm{mRNA}_{\mathrm{A}}$\tabularnewline
        $\mathrm{mRNA}_{\mathrm{A}}+\mathrm{A}$  & $\rightharpoonup$  & $\mathrm{mRNA}_{\mathrm{A}}$  & $k_{19}=0$  & artificial fusion\tabularnewline
        $\mathrm{mRNA}_{\mathrm{B}}+\mathrm{B}$  & $\rightharpoonup$  & $\mathrm{mRNA}_{\mathrm{B}}$  & $k_{20}=0$  & artificial fusion\tabularnewline
        \bottomrule
    \end{tabular}\caption{\label{tab:reaction-library-lv}Full set of ansatz reactions $\Theta$ used in Sec.~\ref{sec:lv} for the Lotka-Volterra system. The given rate constants define the ground truth reaction model.}
\end{table*}
